---
title: "140.615.HW.02.Jin.Vincent"
author: "Vincent Jin"
date: "2023-02-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 02
## Vincent Jin

## 1. Suppose that 1/100,000 bacterial cells contain a mutation providing resistance to substance A. Suppose I create a bunch of plates with 200,000 cells per plate. Let X be the number of cells on a plate that are resistant to substance A.

(a) What is the mean number of cells in a plate that are resistant to A?

***Answer***

``` {r Q1a}
res_p <- 1 / 100000
total_c <- 200000
mean_x <- res_p * total_c
print(mean_x)
```

The mean number of cells in a plate that are resistant to A is 2.

(b) What is SD(X)?

***Answer***

As the number of cells in a plate that are resistant to A follows a poisson distribution, the variance will be equal to the mean, which means the variance is equal to 2.

``` {r Q1h}
sd_x <- sqrt(mean_x)
print(round(sd_x, 2))
```

Therefore, SD(X) equals to the square root of mean of x, which is 1.41.

(c) Calculate Pr(X = 0).

***Answer***

```{r}
Pr0 <- dpois(0, mean_x)
print(round(Pr0, 2))
```
The probability of X = 0 is 0.14.

(d) Calculate Pr(X = 5).

***Answer***

```{r}
Pr5 <- dpois(5, mean_x)
print(round(Pr5, 2))
```
The probability of X = 5 is 0.04. 

(e) Calculate Pr(X > 2).

***Answer***

```{r}
Pr2 <- 1 - ppois(2, mean_x)
print(round(Pr2, 2))
```
The probability of X > 2 is 0.32.


## 2. Suppose Y is a random variable with E(Y) = 30 and SD(Y) = 5.

(a) Let Z = (Y – 30)/5. Calculate E(Z) and SD(Z).

***Answer***

$$ Z = (Y - 30) / 5 = Y / 5 - 6 $$

``` {r Q2a}
E_Y <- 30
SD_Y <- 5
E_Z <- (E_Y - 30) / 5
SD_Z <- 1 / 5 * SD_Y
print(E_Z) 
print(SD_Z)
```
Therefore, we can see that:
E(Z) = 0
SD(Z) = 1

(b) Let X = – Y. Calculate E(X) and SD(X).

***Answer***

``` {r Q2b}
E_X <- -1 * E_Y
SD_X <- abs(-1) * SD_Y
print(E_X) 
print(SD_X)
```
Therefore, we can see that:
E(X) = -30
SD(X) = 5

(c) Let R = 5 + Y/3. Calculate E(R) and SD(R).

***Answer***

``` {r Q2c}
E_R <- 5 + 1/3 * E_Y
SD_R <- 1/3 * SD_Y
print(E_R) 
print(SD_R)
```
Therefore, we can see that:
E(R) = 15
SD(R) = 1.67


## 3. Below is the joint probability distribution of viral infection and strain in a population of rats.

Strain          A       B       C
Infected        0.25    0.19    0.03
Not infected    0.19    0.17    0.17

(a)  If we randomly sample a rat, what is the probability it is infected with the virus?

***Answer***

``` {r Q3a}
rows <- c("Infected", "Not Infected")
cols <- c("A", "B", "C")
joint_p <- matrix(c(0.25, 0.19, 0.03, 0.19, 0.17, 0.17), nrow = 2, byrow = TRUE, dimnames = list(rows, cols))
P_inf <- as.numeric(joint_p[1, 1]) + as.numeric(joint_p[1, 2]) + as.numeric(joint_p[1, 3])
print(P_inf)
```
Therefore, the probability of randomly sample an infected rat is 0.47.

(b)  If we randomly sample a rat, what is the probability it is of strain A?

***Answer***

``` {r Q3b}
P_A <- joint_p[1, 1] + joint_p[2, 1]
print(P_A)
```
Therefore, the probability of randomly sample an strain A rat is 0.44.

(c) What is the strain distribution among infected rats?

***Answer***

``` {r Q3c}
P_Ainf <- round(joint_p[1, 1] / P_inf, 2)
P_Binf <- round(joint_p[1, 2] / P_inf, 2)
P_Cinf <- round(joint_p[1, 3] / P_inf, 2)
inf <- matrix(c(P_Ainf, P_Binf, P_Cinf), nrow = 1, byrow = TRUE, dimnames = list("Infected", cols))
print(inf)
```
Therefore, among infected rats, the probability of being strain A, B, and C is 0.53, 0.40, 0.06, respectively.

(d) Are the random variables strain and infection independent?

***Answer***

``` {r Q3d}
Ind_TF <- (P_inf * P_A) == joint_p[1, 1]
print(Ind_TF)
```
Since the probability of being infected and is strain A does not equal to the production of being infected and probability of being strain A, the random variables strain and infection are not independent.


## 4. Assume blood-glucose levels in a population of adult women are normally distributed with mean 90 mg/dL and standard deviation 13 mg/dL. What percentage of individuals would be have ’abnormal’
glucose levels, and what is the normal range of glucose levels (in units of mg/dL), if ’abnormal’ was defined as glucose levels outside of 2 standard deviations of the mean?

***Answer***

``` {r Q4}
bg_mean <- 90
bg_sd <- 13
normal_upper <- bg_mean + 2 * bg_sd
normal_lower <- bg_mean - 2 * bg_sd
print(c(normal_lower, normal_upper))
```
Since there will be 95% of population within 2 standard deviation from the mean, if abnormal was defined as gluciose levels outside of 2 standard deviations of the mean, so that there will be 1 - 0.95 = 0.05 = 5% of adult women would be having abnormal glucose levels. The normal range of glucose level is 64 mg/dL - 116 mg/dL.


## 5. Suppose we have 100 independent draws from a normal distribution with population mean 20 and population standard deviation 5.

(a) What is the chance that the sample mean is within 0.1 units of the population mean?

***Answer***

Since we sampled from a population, we can know that in this sample, the sample mean will be equal to the population mean, and the sample mean standard deviation will be the standard error of $s/\sqrt{n}$.
``` {r Q5a}
se_mean <- 5 / sqrt(100)
P_0.1 <- pnorm(20 + 0.1, 20, se_mean) - pnorm(20 - 0.1, 20, se_mean)
print(round(P_0.1, 2))
```
The chance that the sample mean is within 0,1 units of the population mean 16%.

(b) What is the chance that the sample mean exceeds the population mean by at least 0.25 units?

***Answer***

``` {r Q5b}
P_ex0.25 <- 1 - pnorm(20 + 0.25, 20, se_mean)
print(round(P_ex0.25, 2))
```
The chance that the sample mean exceeds the population mean by at least 0.25 units is 31%.

